A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-08-13 appendix.tex

\section{Variational principle for locally isothermal gas
  well-coupled to dust}\label{var_prin}

The linearized equations for axisymmetric disturbances are

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
 \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}.
\end{align}  
These equations do not assume the radially-local approximation used in
the main text.

From the linearized meridional momentum equations, we find 
\begin{align}
  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV = \int\left( \rho
  \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
  v_r^*  + \ii\sigma \rho \dd \bm{F}\cdot\dd\bm{v}^*
  \right)dV,
\end{align}
where the integral is taken over the volume of the fluid. Integrating
by parts and ignoring surface integrals, the last term is
\begin{align}
  \int \ii\sigma\rho \dd \bm{F}\cdot\dd\bm{v}^*dV &=\int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot \nabla P - \ii\sigma \dd
  \bm{v}^*\cdot\nabla \dd P  \right)dV\notag\\
&= \int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot\nabla P + \ii\sigma \dd
 P  \nabla\cdot \dd \bm{v}^*  \right)dV\notag\\
 & = \int\left[
   \left(
   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
   \bm{v} \cdot \nabla \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}
   \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
   P  \nabla\cdot \dd \bm{v}^*
   \right]dV\notag\\
 & = \int\left[
   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s\right) 
   \right]dV\notag\\
 & = \int\left\{
   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
   \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}  \right]\nabla\cdot\left(P\dd\bm{v}^*\right)
   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s \right)
   \right\}dV \notag\\
   &=\int\left[
     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
     \right) -
     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
     \right]dV,
\end{align}
where we recall $s\equiv\ln{(P/\rho)}$. Hence,
\begin{align}\label{integral_ex}
  &\sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
&=  \int\left\{
  \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p r}\right)
  + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p z}\right)
   + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
  \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p z}\right) 
  + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p r}\right)\right. \notag\\
&\phantom{==\int}\left.+
  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
  \right\}dV -\int
  P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV
  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV.  
\end{align}
Note that the coefficients of the mixed terms are equal owing to the
equilibrium state.

@@

\section{Dispersion relation for unstratified disks}
Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
0$. The background $\p_rP/\rho$ and $\tepsilon$ are constant. 
Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
$\p_r\to \ii k_x$ when acting on perturbations. Denote $|\bm{k}|^2
\equiv k_x^2 + k_z^2$. Neglect variations in
background when compared with that of perturbations. Linearized
equations are, after eliminating the azimuthal velocity:

\begin{align}
  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}

Eliminate the velocity perturbations to obtain
\begin{align}
  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
  \ii \left[
  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
  2\tepsilon\right)\right]W. \\
   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q. 
\end{align}
Futher eliminate $Q$ and $W$ to obtain the dispersion relation
\begin{align}
  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 \notag\\ 
  &+ \left(1 -
  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
  k_x \tstop F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align}
where the equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
was used. 
If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to obtain the
incompressible limit, and define $\tau_\mathrm{s} \equiv
\tstop/\left(1 - \tepsilon\right)$, we obtain: 
\begin{align*}
\ii \tepsilon \tstop
  \sigma^4 - \sigma^3 - 
  \tau_\mathrm{s} \left[ \ii \tepsilon \kappa^2 + k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 
  + \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma + 
  k_x \tau_\mathrm{s} F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align*}
as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
instability.

>>>>

\section{Variational principle}\label{var_prin}

Here we consider the more general equation of state for the gas with
$P = K \rhog^{\Gamma}$ where $K$ is a prescribed function of position
and $\Gamma$ is a constant. The effective energy equation is then
Eq. \ref{poly_energy}. The linearized equations for axisymmetric
disturbances are

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
%  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
   \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}, 
\end{align}  
where the linearizd pressure force $\delta \bm{F}$ is given in
Appendix \ref{lin_press}. These equations do not assume the
radially-local approximation used in the main text. 
From the linearized meridional momentum equations, we find 
\begin{align}
  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV = \int\left( \rho
  \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
  v_r^*  + \ii\sigma \rho \dd \bm{F}\cdot\dd\bm{v}^*
  \right)dV,
\end{align}
where the integral is taken over the volume of the fluid. Integrating
by parts and ignoring surface integrals, the last term is
\begin{align}
  \int \ii\sigma\rho \dd \bm{F}\cdot\dd\bm{v}^*dV &=\int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot \nabla P - \ii\sigma \dd
  \bm{v}^*\cdot\nabla \dd P  \right)dV\notag\\
&= \int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot\nabla P + \ii\sigma \dd
 P  \nabla\cdot \dd \bm{v}^*  \right)dV\notag\\
 & = \int\left[
%   \left(
%   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
%   \bm{v} \cdot \nabla \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}
%   \right)\dd\bm{v}^*\cdot\nabla P  
   \left(
   \ii\sigma \frac{\dd P}{\Gamma P} - \dd\bm{v}\cdot\nabla s + \frac{\dd
   \bm{v}}{\Gamma} \cdot \nabla \ln{K} +
   \frac{\dd\mathcal{C}}{\Gamma P} \right)\dd\bm{v}^*\cdot\nabla P
   + \ii\sigma \dd P  \nabla\cdot \dd \bm{v}^*
   \right]dV\notag\\
 & = \int\left[
%   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
%   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
%   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s\right) 
%   \right]dV\notag\\
   \ii\sigma\frac{\dd P}{\Gamma P}\left(\bm{\dd v}^*\cdot\nabla P +
   \Gamma P\nabla\cdot\bm{\delta v}^*   \right)
   +\left(\dd\bm{v}^*\cdot\nabla P\right) \left( \frac{\dd\bm{v}}{\Gamma} \cdot \nabla
   \ln{K} + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s\right) 
   \right]dV\notag\\
 & = \int\left\{
%   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
%   \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}  \right]\nabla\cdot\left(P\dd\bm{v}^*\right)
%   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
%   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s \right)
%   \right\}dV \notag\\
 \left[
   \nabla\cdot\dd\bm{v} +
  \frac{\dd\bm{v}}{\Gamma}\cdot\nabla\ln{P} -
  \frac{\dd\bm{v}}{\Gamma}\cdot\nabla\ln{K} -
  \frac{\dd\mathcal{C}}{\Gamma P}. 
  \right]
 \left(\bm{\dd v}^*\cdot\nabla P +
 \Gamma P\nabla\cdot\bm{\delta v}^*   \right)\right. \notag\\
 &\phantom{=\int\left\{\right\}}\left. 
 + \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\frac{\dd\bm{v}}{\Gamma} \cdot \nabla
 \ln{K}
 + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s \right)
 \right\}dV \notag\\
 &=\int\left[
%     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
%     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
%     \right) -
%     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
    \frac{1}{\Gamma P} \Bigl\lvert \bm{\dd v}\cdot\nabla P +
 \Gamma P\nabla\cdot\bm{\delta v}    \Bigr\rvert ^2 - 
     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
     \right) -
      P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{K} + \frac{\dd\mathcal{C}}{P} \right) 
     \right]dV,
\end{align}
where  $s\equiv\ln{(P^{1/\Gamma}/\rho)}$. Hence,
\begin{align}\label{integral_ex}
  &\sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
&=  \int\left\{
  \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p r}\right)
  + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p z}\right)
   + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
  \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p z}\right) 
  + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p r}\right)\right. \notag\\
&\phantom{==\int}\left.+
%  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
 \frac{1}{\Gamma P}\Bigl\lvert \bm{\dd v}\cdot\nabla P +
 \Gamma P\nabla\cdot\bm{\delta v}     \Bigr\rvert^2
  \right\}dV 
%  P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV
-\int    P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{K}\right)dV
  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV.  
\end{align}
Note that the coefficient of $\delta v_z\delta v_r^*$ and $\delta
v_z^*\delta v_r$ are equal owing to the
equilibrium state (Eq. \ref{vshear}). 
The special case $\Gamma=1,\, K = c_s^2$ gives
Eq. \ref{int_rel}. A radially-local, vertically-global, viscous version of
Eq. \ref{integral_ex} is given by \cite{latter06}. See also
\cite{kato78} .

@@

\section{Dispersion relation for unstratified disks}\label{compressible_streaming}
Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
0$. The background $\p_rP/\rho$ and $\tepsilon$ are constant. 
Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
$\p_r\to \ii k_x$ when acting on perturbations. Denote $|\bm{k}|^2
\equiv k_x^2 + k_z^2$. Neglect variations in
background when compared with that of perturbations. Linearized
equations are, after eliminating the azimuthal velocity:

\begin{align}
  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}
We note that the incompressible condition used by \cite{jacquet11} is
obtained by setting $\zeta$ to zero.

Eliminate the velocity perturbations to obtain
\begin{align}
  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
  \ii \left[
  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
  2\tepsilon\right)\right]W. \\
   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q. 
\end{align}
Futher eliminate $Q$ and $W$ to obtain the dispersion relation
\begin{align}
  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 \notag\\ 
  &+ \left(1 -
  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
  k_x \tstop F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align}
where the equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
was used. 
If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to obtain the
incompressible limit, and define $\tau_\mathrm{s} \equiv
\tstop/\left(1 - \tepsilon\right)$, we obtain: 
\begin{align*}
\ii \tepsilon \tstop
  \sigma^4 - \sigma^3 - 
  \tau_\mathrm{s} \left[ \ii \tepsilon \kappa^2 + k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 
  + \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma + 
  k_x \tau_\mathrm{s} F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align*}
as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
instability. If $|\sigma| = O(\tau_\mathrm{s})$ then the quartic term
is small and may be neglected. In that case we obtain the cubic
dispersion relation of \citet{youdin05a}.

### 2016-08-13 model.tex

\begin{align}\label{vshear_split}
  \frac{\left|r\p_z\Omega\right|_{\text{
        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
        temp. gradient}}} \sim
  \epsilon \frac{\mathrm{max}\left(\delta^2,
    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
\end{align}
where $\delta\equiv \Htilde/\Hgas$. 
Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
vertical shear due to variations in the dust-to-gas ratio is at 
most comparable to that due to the radial temperature gradient. 
This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so
the above ratio $\sim 1/|q|$). Otherwise, vertical shear is 
predominantly associated with  $\p_rc_s^2$.

\subsection{Vertical buoyancy}\label{vbuoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thin dust layers such that $H_\epsilon \ll Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

>>>>

\begin{align}\label{vshear_split}
  \frac{\left|r\p_z\Omega\right|_{\text{
        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
        temp. gradient}}} \sim
  \epsilon \frac{\mathrm{max}\left(\delta^2,
    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
\end{align}
where $\delta\equiv \Htilde/\Hgas$. 
Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
vertical shear due to variations in the dust-to-gas ratio can be at
most comparable to that due to the radial temperature gradient. 
This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so 
the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
with $\p_rc_s^2$.

\subsection{Vertical buoyancy}\label{vbuoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
